---
title: "m15_202111_baseline -150 to 50, 200-300 ms difference wave"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages and define functions
This section load the packages `knitr`, `markdown`, `ez`, `stringr` `readr`, `tidyr`, and `dplyr`.
It also define a function to compute the standard error of the mean and to calculate the mean, standard deviation and standard error for each condition.
## Load Packages

```{r load_packages, echo=FALSE, message=FALSE}
require(knitr)
require(markdown)
require(ez)
require(stringr)
require(readr)
require(tidyr)
require(dplyr)
```

## Function to calculate the standard error of the mean 

```{r sem}
sem = function(x)
{
  sqrt(var(x)/length(x))

}
```


## Function to calculate the mean, the standard deviation and the standard error for each condition

`data` : a data frame
`varname` : the name of a column containing the variable to be summariezed
`groupnames` : vector of column names to be used as grouping variables

```{r data_summary}

data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE),
      sem = sd(x[[col]])/sqrt(length(x)))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
 return(data_sum)
}
```


# Analyse Affix Frequency

## Read in and format the data
Then filter into two datasets, one with 2 Relatedness Factors (related, unrelated) and 2 Productivity Factors (high, low) and another with just one factors—Priming Effects for High adn Low productivity calculated by subtracting Related scores from Unrelated.

```{r read_data, message=FALSE}
m15_200_300_afx <- read_csv("M15_afxfrq_200_300_bsl_150_50.csv")
m15_diff_afx <-  filter(m15_200_300_afx, binlabel == "Priming_High" | binlabel == "Priming_Low")
m15_2by2_afx <-  filter(m15_200_300_afx, binlabel != "Priming_High" & binlabel != "Priming_Low")
```

Add factors *relatedness* and *productivity* for the 2-factor dataframe by separating 'binlabel' variable. Recodes the difference wave dataframe by removing the "Priming" part of the binlabel.

```{r add_rel_prod}
m15_2by2_afx <-separate(m15_2by2_afx, binlabel, into = c("relatedness", "productivity"), sep = "_")

m15_diff_afx$productivity <- ifelse(m15_diff_afx$binlabel == "Priming_Low", "Low", "High")
m15_diff_afx$binlabel <- NULL # removes binlabel column; no longer needed
```

Separate electrode labels into multiple factors based on *anteriority* and *laterality*. `tidyr::separate` makes separating columns simple by allowing you to pass an integer index of split position, including negatively indexed from the end of the string.

```{r add_ant_lat}
m15_2by2_afx <- m15_2by2_afx %>% 
  separate(chlabel, into = c('anteriority', 'laterality'), sep = -1, convert = TRUE)

m15_2by2_afx <- m15_2by2_afx %>%                               
  mutate(laterality = replace(laterality, laterality == "Z", 0))  # Replacing "Z" value with 0

#Extract 5 x 3 matrix for analysis (F3 to P4)

m15_2by2_afx_subset <-  filter(m15_2by2_afx, laterality == 0 & anteriority!= "O" | 
                                 laterality == 3 | laterality == 4)
```